
---
title: 'BiH: Voting behavior of MPs in House of Representatives'
author: ro
date: '2019-11-20'
slug: bih-voting-behavior-of-mps-in-house-of-representatives
categories: []
draft: true
tags:
  - Bosnia
description: ''
---

# Context
A while ago I noticed that the website of Bosnia's parliament now also provides the voting records for each separate vote taken in both houses. These records also include the voting behavior for each individual MP. While the dominant role of the ethnic cleavage in Bosnian politics has been a constant theme and continuously emphasized for the last 25 years (to an extent which frequently blanks out other salient issues), I hadn't seen any  empirical record on MP's voting behavior which could cooroberate the cleavage's prevalence in the legislative arena (one insightful exception is Birgit Kunrath's 2011 [paper](https://www.cambridge.org/core/journals/nationalities-papers/article/of-veto-players-and-entityvoting-institutional-gridlock-in-the-bosnian-reform-process/96BB80A5191E12B40F501841546C151D){target="_blank"}). 

This post digs into the now available records for the 2014-2018 parliament and seeks to shed some light on the legislative process. The first part presents the output, the second highlights the most relevant steps in R to produce them. The entire code for the analysis is again available on my [github account](http://www.github.com/werkstattcodes){target="_blank"}. 

As always, if you see any glaring error, feel free to let me know (best via twitter DM). If you think the blog is interesting and feel like quoting it, please refer to it as Roland Schmidt, "", http:://werk.statt.codes/BiH_MP_voting_2014_2018, 12 December 2019.

For those working in R, the blog might be particularly interesting when it comes to retrieving the acutal data. As so often with empiricalresearch the challenges has been less the analysis of the data itself, but  'liberating' it from the format it is provided. In this case, the the voting records for both the House of Peoples (upper chamber) and the House of Representative (lower chamber) are provided in pdfs, such as this one or this one. Hence, there's been some hefty OCR-ing involved thanks to the powerfull pdf_tools package by ... .


# Results

```{r include=FALSE}
# setup -------------------------------------------------------------------
library(tidyverse)
library(paletteer)
library(glue)
library(extrafont)
library(paletteer)
loadfonts(device = "win", quiet = T)
library(hrbrthemes)
library(paletteer)
library(ggtext)
library(scales)
library(patchwork)
library(padr)
library(here)

wdr <- getwd() 

my_caption <- c("Data: parlament.ba", "Analysis: Roland Schmidt | @zoowalk | <span style='color:black'>**werk.statt.codes**</span>")

df_ocr_results_unnested <- readr::read_csv2(here("blog_data", "Bosnia_MP_voting_behavior", "df_ocr_results_unnested.csv"))
  
df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate_if(is.character, as.factor) 
  
#parse_date_time order allows to specify different formats and their priority
#could be moved to other R script
df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate(session_date=as.character(session_date)) %>% 
  mutate(session_date=lubridate::parse_date_time(session_date,
                                                 orders=c("dmy","ymd", "mdy")) %>% 
           lubridate::as_date(.)) 

#order factor of vote
df_ocr_results_unnested <- df_ocr_results_unnested %>% 
  mutate(vote=forcats::fct_infreq(vote))


voting_colors <- c("no"="red",
                   "no vote" = "grey",
                   "not present" = "brown",
                   "yes" = "green",
                   "restrained"="orange")


# _ ----------------------------------------------------------------------

# FILTER ------------------------------------------------------------------

#identify start of new legislative session
date_new_parliament <- df_ocr_results_unnested %>% 
  filter(str_detect(delegate_name_2, "^Ahmetovic")) %>% 
  filter(session_date==min(session_date)) %>% 
  distinct(session_date) %>% 
  pull(session_date)


df_ocr_results_unnested_truncated <- df_ocr_results_unnested %>% 
  filter(session_date >= as.Date(date_new_parliament)) %>% 
  mutate(party=forcats::fct_drop(party)) #remove empty party factor levels

```

## Number of bills considered in both chambers and outcome.

Overall, during the 2014 to 2018 parliament session 163 bills were considered. Out of this total, only 37 % were eventually passed and published to become law. 38 % of all bills were rejected in at least one of both houses.

```{r echo=FALSE, fig.height=3, fig.width=10, message=FALSE}
df_details_all_status_wide <- readr::read_csv2(here::here("blog_data","Bosnia_MP_voting_behavior","df_details_all_status_wide.csv")) %>% 
    mutate_at(vars(contains("status")), as.factor)

#standardize colors across plots
colors_outcome <- paletteer_d("ggsci::default_jco", n = 10) %>% as.character() 
names(colors_outcome) <- unique(c(levels(df_details_all_status_wide$HoP_status), 
                                  levels(df_details_all_status_wide$HoR_status), 
                                  levels(df_details_all_status_wide$BiH_Parl_final_status)))


df_laws_outcome <- df_details_all_status_wide %>% 
  group_by(BiH_Parl_final_status) %>% 
  summarise(n_obs=n()) %>% 
  ungroup() %>% 
  mutate(n_rel=n_obs/sum(n_obs)) 

df_laws_outcome %>% 
  ggplot()+
  labs(title="Outcome of bills",
       subtitle="Outcome of bills considered during legislative session 2014-2018; Total number: 163.",
       caption=my_caption)+
  geom_bar(aes(x=reorder(BiH_Parl_final_status, n_rel),
               y=n_obs,
               fill=BiH_Parl_final_status),
           stat="identity",
           position=position_dodge())+
  geom_text(aes(x=reorder(BiH_Parl_final_status, n_rel),
                y=80,
                label=paste0(n_obs, " (",
                  percent(n_rel, accuracy = 1),
                  ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = 1,
           # group = name,
            stat="identity")+
  scale_fill_manual(values=colors_outcome)+
  #scale_fill_paletteer_d("ggsci::default_jama")+
  scale_y_continuous(breaks=seq(0,60, 20),
                     expand=expansion(mult=c(0)))+
  hrbrthemes::theme_ipsum_rc() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title=element_blank(),
        plot.margin=margin(l=0, t=0.5, b=0.5, unit="cm"),
        plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
        plot.title.position = "plot",
        plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.caption.position = "plot",
        plot.caption = element_markdown(color = "grey30", hjust = c(0, 1)))+
  coord_flip()



```
##by houses
Looking separately at each of the two houses provides a more detailed picture. In the lower chamber, the House of Representatives, overall 153 bills were considered out of which a bit more than 50 % were approved. In the upper chamber, the House of Representatives, which comprises 5 delegates of each constituent people (Bosniak, Croats, Serbs), 56.7 % of all considered bills were passed. 

```{r echo=FALSE, fig.height=3, fig.width=10}
#in which house are bills rejected

df_outcome_chambers <- df_details_all_status_wide %>% 
  select(contains("status"), -contains("final")) %>% 
  pivot_longer(cols=contains("status"), names_to="house", values_to="outcome") %>% 
  filter(outcome!="missing") %>%
  mutate(outcome=fct_drop(outcome, only="missing")) %>%
  mutate(outcome=fct_infreq(outcome) %>% fct_rev) %>% 
  group_by(house, outcome, .drop=F) %>% 
  summarise(n_obs=n()) %>% 
  group_by(house) %>% 
  mutate(n_rel=n_obs/sum(n_obs)) %>% 
  arrange(desc(n_obs), .by_group=T) %>% 
  mutate(cumsum_n_obs=cumsum(n_obs)) %>% 
  ungroup() %>% 
  mutate(outcome_label = case_when(house=="HoR_status" ~ paste0(outcome, "\n(", n_obs, ")"),
                                   house=="HoP_status" ~ paste0(outcome, "(", n_obs, ")")))

plot_df_outcome_chambers <- df_outcome_chambers %>% 
  # mutate(outcome=fct_drop(outcome, only="missing")) %>% 
  # filter(outcome!="missing") %>% 
  group_split(house) %>% 
  map(~ggplot(.)+
  geom_bar(aes(x=1,
               y=n_obs,
               fill=outcome,
               group=outcome),
           stat="identity",
           position = position_stack())+
  geom_text(aes(x=1,
                y = n_obs,
                label=ifelse(n_obs>0,
                             n_obs,
                             ""),
                group=outcome),
            family = "Roboto Condensed",
            color = "white",
            position = position_stack(vjust= 0.5))+
  geom_text(aes(x=1,
                y = n_obs,
                label=ifelse(n_obs>10,
                             paste0("\n", scales::percent(n_rel)),
                             ""),
                group=outcome),
            family = "Roboto Condensed",
            color = "white",
            position = position_stack(vjust= 0.5))+  
  labs(y="Number of bills",
       title=ifelse(unique(.$house)=="HoP_status", 
                    "House of Peoples",
                    "House of Representatives"))+
      #subtitle="Bills considered in 2014 - 2018 legislative period.",
      # caption=my_caption)+
  scale_fill_manual(values=colors_outcome)+
  # scale_y_continuous(expand=expansion(mult=c(0,0.1)),
  #                    breaks=df_outcome_chambers %>% 
  #                      filter(house==.$house) %>% 
  #                      pull(cumsum_n_obs))+
  scale_y_continuous(expand=expansion(mult=c(0,0.1)),
                     limits=c(0, max(df_outcome_chambers$cumsum_n_obs)),
                     breaks=.$cumsum_n_obs,
                     labels=ifelse(.$n_obs<10, 
                                   ifelse(.$cumsum_n_obs==max(.$cumsum_n_obs), 
                                          unique(max(.$cumsum_n_obs)), ""),
                                          .$cumsum_n_obs))+
  coord_flip()+
  # lemon::facet_rep_wrap(vars(house),
  #                       repeat.tick.labels = F,
  #                       nrow=2,
  #                       labeller = labeller(house=c("HoP_status"="House of Peoples",
  #                                                "HoR_status"="House of Representatives")))+
  theme_ipsum_rc()+
  theme(panel.grid = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        plot.margin=margin(b=0, t=0, l=0, r=0, unit="cm"),
        legend.title = element_blank(),
        legend.position = "bottom")+
  guides(fill=guide_legend(reverse = T, 
                           nrow=1)))

(patchwork::wrap_plots(plot_df_outcome_chambers))+
  plot_layout(ncol=1, guides="collect") &
  theme(legend.position = "bottom",
        legend.justification = "left")

```
##intersection of houses
If we intersect the developments for each law in each chamber, we get a clearer were bills were actually killed. The table/heatmap below intends to show this. The x-axis shows a bill's outcome in the House of Representative, the y-axis the same bill outcome in House of Peoples (if a bill considered only in one chamber, the respective status in the other chamber is 'missing').

Out of the 67 bills adopted in the House of Peoples and 77 in the House of Representatives, 60 were acutally considered in both chamber and eventully passed. This With 37.0 % of all bills it's the largest category in our table. The second largest category comprises 27 bills (16.7 % of all bills) which were rejected in the House of Representatives and never made it to the House of Peoples. The third largest category contains 12 bills (7.4 % of all bills) which were adopted in the lower house but rejected in the upper chamber. 


```{r echo=FALSE, fig.width=10, message=FALSE}
df_outcome_chambers_combined <- df_details_all_status_wide %>% 
  mutate_at(vars(contains("status")), as.factor) %>% 
  group_by_at(vars(contains("status"), -contains("final")),
              .drop=F) %>% 
  summarise(n_obs=n()) %>% 
  ungroup() %>% 
  mutate(n_rel=n_obs/sum(n_obs)) %>% 
  group_by(HoR_status) %>% 
  mutate(HoR_outcome_label=sum(n_obs) %>% paste0(HoR_status, " (", ., ")")) %>% 
  group_by(HoP_status) %>% 
  mutate(HoP_outcome_label=sum(n_obs) %>% paste0(HoP_status, " (", ., ")")) %>% 
  ungroup() %>%  #ungroup before converting to factor
  mutate_at(vars(contains("label")), as_factor)

#to duplicate axis labels scale has to be numeric; convert to numeric, but
#us character labels;

plot_df_outcome_chambers_combined <- df_outcome_chambers_combined  %>% 
  ggplot()+
  labs(title="Outcome of bills by house",
       x="final status in House of Representatives",
       y=str_wrap("final status in House of Peoples", 20),
       caption=my_caption)+
  geom_tile(aes(x=as.numeric(HoR_outcome_label),
                y=as.numeric(HoP_outcome_label),
                fill=n_obs))+
  geom_text(data=. %>% filter(n_obs!=0),
              aes(x=as.numeric(HoR_outcome_label),
                y=as.numeric(HoP_outcome_label),
                label=paste0(n_obs, " (",
                             percent(n_rel, accuracy = 0.1),
                             ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = .5,
            # group = name,
            stat="identity")+
  scale_fill_gradient(low="white", high="steelblue",
                      na.value="grey")+
  scale_x_continuous(labels=levels(df_outcome_chambers_combined$HoR_outcome_label) %>% map_chr(., label_wrap(15)),
                     breaks=seq(1, length(unique(df_outcome_chambers_combined$HoR_outcome_label))),
                     sec.axis = dup_axis())+
  scale_y_continuous(labels=levels(df_outcome_chambers_combined$HoP_outcome_label) %>% map_chr(., label_wrap(20)),
                     breaks=seq(1, length(unique(df_outcome_chambers_combined$HoP_outcome_label))),
                     sec.axis = dup_axis(),
                     trans="reverse")+
  theme_ipsum_rc()+
  theme(axis.text.x.top=element_blank(),
        axis.title.x.top = element_text(hjust=0, size=11),
        axis.title.x.bottom = element_blank(),
        axis.text.y.left=element_blank(),
        axis.title.y.left = element_text(hjust=0, size=11),
        axis.title.y.right = element_blank(),
        #axis.text.x.bottom=element_text(angle=0, hjust=.5),
        panel.grid = element_blank(),
        legend.position = "none",
        panel.border = element_rect(color="grey30", fill=NA),
        axis.title.y = element_text(angle=0),
        plot.title = element_text(size = 12, face="bold.italic", margin=margin(b=0, unit="cm")),
        plot.title.position = "panel",
        plot.subtitle = element_text(size = 12, color = "grey30"),
        plot.caption.position = "panel",
        plot.caption = element_markdown(color = "grey30", hjust = c(0, 1)))


plot_df_outcome_chambers_combined


```

## time line
How was legislative activity distributed over time? As the graph below shows, both chamber were most active in the first half of 2016 and the end of 2017. I am not sure exactly what caused this bout of activity, but it's quite distinct if contrasted e.g. wtih 2015 or the early parts of 2017.

```{r echo=FALSE, fig.width=10}
#number of votes per week
#use floor_date to aggregate witin a week
#use pdr to thicken to weeks, then pad for missing weeks, then plot;

df_ocr_results_unnested %>% 
  mutate(month=lubridate::floor_date(session_date, unit="month")) %>% 
  group_by(house, month) %>% 
  summarise(n_votes=n_distinct(record_id)) %>%
  pad(., interval="month",
      start_val=min(.$month),
      end_val=max(.$month)
      ) %>% 
  padr::fill_by_value() %>% 
  mutate(year=lubridate::year(month)) %>% 
  ggplot()+
  labs(subtitle="Number of items voted on per month.",
       title="Legislative activity",
       caption=my_caption)+
  geom_bar(aes(y=n_votes,
               x=month),
           fill="steelblue",
           stat="identity")+
  scale_x_date(date_breaks = "1 month",
               labels=scales::label_date(format="%b"))+  #prints year only when year changes
  facet_grid(house~year,
             scales="free_x",
             space="free_x",
             shrink=F)+
  theme_ipsum_rc()+
  theme(axis.text.x = element_text(angle=90, size=10),
        axis.title = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.caption = element_markdown(hjust=c(0,1)),
        panel.spacing.x = unit(0, "cm"))


#check session number and dates

df_sessions <- df_ocr_results_unnested %>% 
  select(house, session_date, session_no) %>% 
  distinct() %>% 
  arrange(house, session_date) %>% 
  mutate(new_parlament=case_when(session_no<lag(session_no) ~ "start",
                                 TRUE ~ NA_character_)) 

```

## tile graph 
So far the data presented was limited to bills as the unit of analysis. Now let's dig a bit deeper and see how individual MPs voted. The graph below provides the results for every vote of every MP, by party affiliation. Since a bill undergoes several readings before it can be become a law/is forwarded to the other chamber, there are much more votes than bills. Overall, MPs in the House of Representative casted their votes in total in 451 ballots.

In the graphs below each vertical line (small rectangle) represents an MP's vote in one ballot (hence the x axis is 451 small rectangles long). According to the ballot records, there are five distinct options: an MP can cast a 1) yes, or 2) no vote; 3) be not present during the ballot; 4) is restrained, or 5) simply does not cast a vote.

```{r echo=FALSE, fig.height=10, fig.width=10}
# tile graph --------------------------------------------------------------
library(ggforce)
library(ggiraph)

n_record_ids <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  pull(record_id) %>% 
  n_distinct()

plots_tile <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  ggplot()+
  labs(x="item voted at",
       y="MP",
       title="Vote by MP and item",
       subtitel=my_caption)+
  geom_tile(aes(x=reorder(record_id, session_date),
                y=fct_rev(delegate_name_family),
                                     fill=vote))+
 # scale_y_discrete(label=delegate_name_2)+
  scale_fill_manual(values=voting_colors)+
  theme_ipsum_rc() +
  theme(axis.text.x=element_blank(),
        panel.grid = element_blank(),
        legend.position = "bottom",
        legend.justification = "right")+
  facet_wrap(vars(party),
             scales="free_y")

plots_tile

```

Even without going into numeric details, I find the graph rather informative from a few persepectives. While with most parties/MPs the yes votes (green) dominate, the SNSD is a glaring aberrations. Milorad Dodik's SNSD clearly stands out as the party which says 'no' (to paraphrase Ian Paisley) or simply doesn't show up when it comes to cast a vote (and hence may frustrate the quorum necessary for e.g. a bill to pass).
 

## MP ranking per vote category 

The graph below presents each MPs voting behavior per voting category as % of an MPs overall votes. In other words, the graph reveals how often an MP votes for e.g. yes, no etc. as % of her overall votes casted over the period of the 2014 to 2018 parliament. Party affiliation is presented next to each MPs name. Furthermore, to highlight discrepancies between representatives originating (in terms of mandate) from the Federation or the RS, the bars are coloured differently.

```{r echo=FALSE, fig.height=25, fig.width=10, message=FALSE, warning=FALSE}
library(tidytext)
library(lemon)

df_votes_delegates  <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  group_by(entity, party, delegate_name_2, delegate_name_family, vote) %>% 
  summarise(n_votes_cat=n()) %>% 
  group_by(party, delegate_name_2, delegate_name_family) %>% 
  mutate(n_votes_total=sum(n_votes_cat)) %>% 
  mutate(n_votes_cat_rel=n_votes_cat/n_votes_total) %>% 
  ungroup()

df_votes_delegates %>% 
  mutate(delegate_name_3=str_extract(delegate_name_2, "[:alpha:]+,\\s[:alpha:]?")) %>% 
  ggplot()+
  labs(title="MP's votes by category (in %)",
       subtitle ="In how many % of all her votes did the MP vote/was.... Absolute number in category and total in brackets",
       caption=my_caption)+
  geom_bar(aes(x=reorder_within(paste0(delegate_name_3, " (", party, ")"), n_votes_cat_rel, vote),
               y=n_votes_cat_rel,
               fill=entity),
           stat="identity",
           position=position_stack())+
  geom_text(aes(x=reorder_within(paste0(delegate_name_3, " (", party, ")"), n_votes_cat_rel, vote),
                y=1.2,
                label=paste0(scales::percent(n_votes_cat_rel, accuracy =0.1), " (", n_votes_cat, "/", n_votes_total, ")")),
            family = "Roboto Condensed",
            color = "grey30",
            hjust = 1,
            stat="identity")+
  scale_fill_paletteer_d("ggsci::default_jama")+
  scale_x_reordered()+
  scale_y_continuous(minor_breaks = NULL,
                     labels=scales::label_percent(accuracy=1),
                     breaks=seq(0,0.8,.1))+
  theme_ipsum_rc()+
  theme(axis.text.x = element_text(angle=90),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.spacing = unit(0, "cm"),
        strip.text = element_text(face="bold"),
        plot.title.position = "plot",
        plot.caption.position = "plot",
        plot.caption = element_markdown(hjust=c(0,1)),
        axis.title.y = element_blank(),
        axis.title.x = element_text(hjust=0),
        legend.position = "none")+
  lemon::facet_rep_wrap(~vote,
                        repeat.tick.labels = T,
                        ncol=2,
                        scales="free_y")+
  coord_flip()
  
```
## Party differences
What becomes pretty quickly clear is that the 


```{r echo=FALSE, fig.height=20, fig.width=10, warning=FALSE}

colors_ethnicity=c("Bosniak"="green",
                   "Croat"="blue",
                   "Serb"="red",
                   "civic"="orange",
                   "independent"="grey")

df_delegate_voting <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  mutate(ethnicity=forcats::fct_relevel(ethnicity, "Bosniak", "Croat", "Serb", "civic", "independent")) %>% 
  group_by(ethnicity, party, delegate_name, vote, .drop=T) %>% 
  summarise(nobs=n()) %>%
  mutate(total.votes=sum(nobs, na.rm = T)) %>% 
  mutate(perc=nobs/total.votes) %>% 
  ungroup()

plot_delegates <- df_delegate_voting %>% 
  group_split(vote) %>% 
  map(~ggplot(., aes(x=reorder_within(party, desc(perc), ethnicity),
                     y=perc,
                     group=party,
                     tooltip=delegate_name,
                     color=ethnicity))+
  geom_jitter_interactive(width=0.03)+
  geom_boxplot()+
  labs(title=paste(unique(.$vote)))+
  hrbrthemes::theme_ipsum_rc()+
  theme(legend.position="none",
        axis.title = element_blank(),
        axis.text.x=element_text(angle=45, 
                                 hjust=0.5),
        strip.text.y = element_text(angle=180, 
                                    vjust=1,
                                    hjust=1),
        plot.margin = margin(t=0, b=0, unit="cm"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.spacing.x = unit(0.3, "cm"),
        panel.spacing.y = unit(0.3, "cm"))+
  scale_y_continuous(labels=scales::label_percent(accuracy=1),
                     minor_breaks = NULL,
                     limits=c(0,NA),
                     position = "right")+
  scale_x_reordered()+
  scale_color_manual(values=colors_ethnicity, na.value="grey")+
  facet_row(~ethnicity,
            # switch = "y",
             scales="free_x",
             space="free"))

wrap_plot_delegats <- wrap_plots(plot_delegates, ncol = 1)

girafe(ggobj=wrap_plot_delegats,
       pointsize=6,
       width_svg=10, height_svg=15)
```

## Entity veto
 
```{r echo=FALSE, fig.height=7, fig.width=10, message=FALSE, warning=FALSE}
# entity voting -----------------------------------------------------------

df_entity_voting <- df_ocr_results_unnested_truncated %>% 
  filter(house=="HoR") %>% 
  group_by(house, entity, law_id, record_id, vote, .drop=F) %>%  #drop to include all vote types
  summarise(votes_n=n()) %>% 
  filter(vote!="not present") %>% #of those present? PENDING!
  mutate(entity_delegates=sum(votes_n)) %>% 
  mutate(votes_rel=votes_n/entity_delegates) %>% 
  filter(vote=="yes") %>% 
  ungroup() %>% 
  arrange(law_id, record_id) %>% 
  #select(entity, record_id, votes_rel) %>% 
  pivot_wider(id_cols=record_id, names_from=entity, values_from = c(entity_delegates, votes_rel, votes_n)) %>% 
  mutate(votes_yes_total_abs = votes_n_FBiH + votes_n_RS) %>% 
  mutate(votes_yes_total_rel = votes_yes_total_abs/(entity_delegates_FBiH+entity_delegates_RS)) %>% 
  mutate(overall_majority = ifelse(votes_yes_total_rel>0.5, "yes", "no")) %>% 
  mutate(RS_veto = ifelse(votes_rel_RS<0.3, "yes", "no"),
         FBiH_veto = ifelse(votes_rel_FBiH<0.3, "yes", "no"))

#one observation wiht NA
# df_entity_voting %>% 
#   filter_at(vars(votes_rel_FBiH, votes_rel_RS, overall_majority),
#             any_vars(is.na(.)))


plot_entity_voting <- df_entity_voting %>% 
  ggplot()+
  labs(x="yes vote-share among FBiH representatives",
       subtitle="Legislative period 2014-2019. House of Representatives only.",
       title="Vote results by entity support in BiH HoR (incl. entity vetoes)",
       caption=my_caption,
       y="yes vote-share among RS representatives")+
  geom_rect(data=. %>% slice(1),
            aes(xmin=1/3, xmax=Inf,
                ymin=-Inf, ymax=1/3),
            alpha=0.3,
            fill="#DF8F44FF")+
  geom_rect(data=. %>% slice(1),
            aes(xmin=-Inf, xmax=1/3,
                ymin=1/3, ymax=1),
            alpha=0.7,
            fill=paletteer_d("ggsci::default_jama")[1])+
  geom_rect(data=. %>% slice(1),
            aes(xmin=-Inf, xmax=1/3,
                ymin=-Inf, ymax=1/3),
            alpha=0.3,
            fill="grey")+
  geom_text(data=. %>% slice(1), aes(x=.02,
                                     y=.2),
            check_overlap = T,
            hjust=0,
            size=4,
            fontface="italic",
            family = "Roboto Condensed",
            color = "grey30",
            label="both entities with vetoes")+
  geom_text(data=. %>% slice(1), aes(x=.02,
                                     y=.7),
            check_overlap = T,
            hjust=0,
            size=4,
            fontface="italic",
            family = "Roboto Condensed",
            color = "white",
            label="Fed entity  veto")+
  geom_text(data=. %>% slice(1), aes(x=.55,
                                     y=.20),
            check_overlap = T,
            hjust=0,
            size=4,
            fontface="italic",
            family = "Roboto Condensed",
            color = "grey30",
            label="RS entity  veto")+
  geom_jitter_interactive(aes(x=votes_rel_FBiH,
                              y=votes_rel_RS,
                              color=overall_majority,
                              shape=overall_majority,
                              tooltip=paste("record id", record_id, "\n",
                                            "total yes:", percent(votes_yes_total_rel, accuracy=0.01),"\n",
                                            "Fed:", percent(votes_rel_FBiH, accuracy=0.01),"\n",
                                            "RS:", percent(votes_rel_RS, accuracy=0.01))),
                          size=2) +
  geom_vline(xintercept = c(1/3))+
  geom_hline(yintercept = c(1/3))+
  theme_ipsum_rc()+
  theme(legend.position = "bottom",
        legend.justification = "right",
        legend.title=element_text(size=10),
        plot.caption=element_markdown(hjust=c(0, 1)),
        axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10),
        axis.title = element_text(size=10),
        panel.grid = element_blank())+
  scale_y_continuous(labels = scales::label_percent(accuracy = 1),
                     minor_breaks = NULL,
                     breaks=c(seq(0, 1, .25), 1/3),
                     expand=expansion(mult=0.02))+
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks=c(seq(0, 1, .25), 1/3),
                     minor_breaks = NULL,
                     expand=expansion(mult=0.02))+
  scale_color_manual(values=c("yes"="seagreen",
                              "no"="firebrick"),
                     na.translate=F,
                     name="Overall majority:")+
  scale_shape_manual(values=c("yes"=19, "no"=1), #not all symbols can be used in ggiraph
                     na.translate=F,
                     name="Overall majority:")

girafe(ggobj=plot_entity_voting,
       pointsize=3,
       width_svg = 10,
       height_svg = 6)


```

## Outcome of votes 
```{r echo=FALSE, fig.height=4, fig.width=10, message=FALSE, warning=FALSE}
# entity voting (bar) --------------------------------------------------


df_overall_majority<- df_entity_voting %>% 
  filter(overall_majority=="no") %>% 
  group_by(overall_majority) %>% 
  summarise(n_obs=n()) %>% 
  mutate(type="no overall majority")

df_entity_vetoes<- df_entity_voting %>% 
  filter(overall_majority=="yes") %>% 
  select(contains("veto")) %>% 
  group_by_all() %>% 
  summarise(n_obs=n()) %>% 
  mutate(type=case_when(RS_veto=="yes" & FBiH_veto=="no" ~ "RS entity veto only",
                        RS_veto=="yes" & FBiH_veto=="yes" ~ "RS and FBiH entity veto",
                        RS_veto=="no" & FBiH_veto=="yes" ~ "FBiH enttiy veto",
                        RS_veto=="no" & FBiH_veto=="no" ~ "passed"))

df_entity_vetoes <- bind_rows(df_overall_majority,
          df_entity_vetoes) %>% 
  select(n_obs, type) %>% 
  mutate(type=fct_rev(fct_relevel(type, "passed", "no overall majority", "RS entity veto only"))) %>% 
  arrange(desc(type)) %>% 
  mutate(n_obs_cum=cumsum(n_obs)) %>% 
  mutate(n_obs_total=sum(n_obs)) %>% 
  mutate(n_obs_rel=n_obs/n_obs_total)

df_entity_vetoes %>% 
  ggplot()+
  labs(y="total number of votes",
       title="Outcome of votes in BiH House of Representatives",
       caption=my_caption)+
  geom_bar(aes(x=1,
               y=n_obs,
               fill=type),
           position=position_stack(),
           stat="identity")+
  geom_text(aes(x=1,
                label=paste(n_obs,
                            "\n\n", 
                            scales::percent(n_obs_rel)),
                y=n_obs),
           position=position_stack(vjust = 0.5),
           family = "Roboto Condensed",
           color = "white",
           stat="identity")+
  coord_flip()+
  hrbrthemes::theme_ipsum_rc()+
  theme(panel.grid = element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        plot.title.position="plot",
        plot.caption=element_markdown(hjust=c(0, 1)),
        plot.caption.position = "plot",
        legend.position = "bottom",
        legend.justification = "right")+
  scale_fill_paletteer_d("ggsci::default_jama",
                         name="Vote result")+
  scale_y_continuous(breaks=df_entity_vetoes$n_obs_cum,
                     expand=expansion(mult=c(0)))+
  guides(fill=guide_legend(reverse = T))

```

# Steps in R
Below the most important steps to obtain the results from above. 

## How to get the data
This has been the most time consumig, but also rewarding part. Two packages were particularly instrumental here: `rvest` to obtain the raw data/pdfs from the parliament's webite and the `pdf_tools` package which is nothing short of magic when it comes to extract data from pdfs.

But first, one step at a time: Essentially, to eventually obtain the pdf documents with the results of the ballots, three steps were needed.

First, get the links to each law.
Second, from there extract the links to each decision
Thrid, extract the data from the ballots' result sheets.



