---
title: 'Number of Submissions: Covid Amendment 2021'
author: rs
date: '2021-01-04'
slug: []
categories: []
tags:
  - Corona virus
  - Austria
description: ''
---

Only very briefly and more a quick sketch: Again a bill seeking to amend Austria's epidemic law. This time concerned parties etc had only three days to file submissions as part of the consultation process. However, the number of submissions reached a record, at least for this legislative period. Note that numbers are based on those published on the parliaments website. 

Next step - how prevalent was the use of a submission template....

<details open>

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyRSS)
library(xml2)
library(rvest)

rss_feed <- "https://www.parlament.gv.at/PAKT/MESN/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP=XXVII&MESN=ME&R_MESN=ME&MIN=ALLE&SUCH=&listeId=102&FBEZ=FP_002"
# df_rss <- tidyRSS::tidyfeed(rss_feed)
df_rss <- xml2::read_xml(rss_feed)
data <- xml_child(df_rss, 1)

df_bill_pages_link <- data %>% 
  xml2::xml_find_all("//link") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="link_single_bill_page") %>% 
  mutate(link_single_bill_page=str_squish(link_single_bill_page)) %>% 
  filter(id>2) #removes first two rows which don't include data on bills

#get title
df_bill_pages_name <- data %>% 
  xml2::xml_find_all("//title") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="title") %>% 
  # mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>% 
  mutate(title=str_trim(title)) %>% 
  filter(id>2) 

#get publication date
df_bill_pages_pub_date <- data %>% 
  xml2::xml_find_all("//pubDate") %>% 
  html_text() %>% 
  enframe(., 
          name="id",
          value="date") %>% 
  mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))

#combine to one dataframe
df_bill <- bind_cols(
  df_bill_pages_name,
  df_bill_pages_pub_date,
  df_bill_pages_link,
) %>% 
  select(-contains("id"))

df_bill <- df_bill %>% 
  mutate(indicator_me=str_extract(link_single_bill_page, regex("(?<=_)\\d*"))) %>% 
  mutate(rss_link_submissions=glue::glue("https://www.parlament.gv.at/PAKT/VHG/XXVII/ME/ME_{indicator_me}/filter.psp?view=RSS&jsMode=&xdocumentUri=&filterJq=&view=&GP=XXVII&ITYP=ME&INR={as.numeric(indicator_me)}&SUCH=&listeId=142&FBEZ=FP_142
"))


fn_get_submissions <- function(rss_link) {

  df_rss <- xml2::read_xml(rss_link)
  data <- xml_child(df_rss, 1)
  
  df_submission_pages_link <- data %>% 
    xml2::xml_find_all("//link") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="link_single_submission_page") %>% 
    mutate(link_single_submission_page=str_squish(link_single_submission_page)) %>% 
    filter(id>2) #removes first two rows which don't include data on submissions
  
  #get title
  df_submission_pages_name <- data %>% 
    xml2::xml_find_all("//title") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="title") %>% 
    mutate(name=str_extract(title, regex("(?<=\\>).*(?=\\<)"))) %>% 
    filter(id>2) %>% 
    select(-title)
  
  #get publication date
  df_submission_pages_pub_date <- data %>% 
    xml2::xml_find_all("//pubDate") %>% 
    html_text() %>% 
    enframe(., 
            name="id",
            value="date") %>% 
    mutate(date=date %>% str_squish() %>% lubridate::dmy_hms(., tz="Europe/Vienna"))
  
  #combine to one dataframe
  bind_cols(
    df_submission_pages_name,
    df_submission_pages_pub_date,
    df_submission_pages_link,
  ) %>% 
    select(-contains("id"))

}

tbl_missing <- tibble(name=NA_character_, date=NA, link_single_submission_page=NA_character_)

df_bill_submission <- df_bill$rss_link_submissions %>% 
  set_names() %>% 
  map_dfr(., purrr::possibly(fn_get_submissions, otherwise = tbl_missing), .id="bill")


df_bill_submission <- df_bill_submission %>% left_join(.,
                                df_bill %>% select(title, rss_link_submissions, date_pub=date),
                                by=c("bill"="rss_link_submissions"))

submission_count <- df_bill_submission %>% 
  group_by(title, date_pub) %>% 
  summarise(n_obs=n()) %>% 
  ungroup() %>% 
  arrange(desc(n_obs))


```

</details>

```{r message=FALSE, warning=FALSE}
reactable::reactable(submission_count)

```

